---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
head(activity)
```

```{r}
names(activity)
```

```{r}
library(lattice)

activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity$interval <- as.factor(activity$interval)
```

## What is mean total number of steps taken per day?

Let's use the aggregate function for this.
```{r}
StepsTaken <- aggregate(steps ~ date, data = activity, sum,na.rm = TRUE)
```

We can view a bar chart for this now
```{r}
hist(StepsTaken$steps,main = "Total number of Steps taken by Day",xlab="Steps",ylab = "Days", col="blue")
```

In order to find the mean, we can just use the mean function on the steps column
```{r}
meanSteps <- mean(StepsTaken$steps,na.rm=TRUE)
medianSteps <- median(StepsTaken$steps,na.rm=TRUE)
```


The Mean is `r round(meanSteps,1)` and the median is `r medianSteps`.

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
timeSeriesPat <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

We plot the time series graph by:
```{r}
plot(row.names(timeSeriesPat), timeSeriesPat, type = "l", xlab = "5-min interval", 
    ylab = "Number of Steps", main = "Average number of steps taken", 
    col = "blue")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

To find that, 
```{r}
MaxStepAvg <- which.max(timeSeriesPat)
IntervalWithMaxStepAvg <- names(MaxStepAvg)
```

The 5-minute interval `r IntervalWithMaxStepAvg` contains the maximum number of steps on average across all the days.

## Inputing missing values



## Are there differences in activity patterns between weekdays and weekends?
