---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
head(activity)
```

```{r}
names(activity)
```

```{r}
library(lattice)

activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity$interval <- as.factor(activity$interval)
```

## What is mean total number of steps taken per day?

Let's use the aggregate function for this.
```{r}
StepsTaken <- aggregate(steps ~ date, data = activity, sum,na.rm = TRUE)
```

We can view a histogram for this now
```{r}
hist(StepsTaken$steps,main = "Total number of Steps taken by Day",xlab="Steps",ylab = "Days", col="blue")
```

In order to find the mean, we can just use the mean function on the steps column
```{r}
meanSteps <- mean(StepsTaken$steps,na.rm=TRUE)
medianSteps <- median(StepsTaken$steps,na.rm=TRUE)
```


The Mean is `r round(meanSteps,1)` and the median is `r medianSteps`.

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
timeSeriesPat <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

We plot the time series graph by:
```{r}
plot(row.names(timeSeriesPat), timeSeriesPat, type = "l", xlab = "5-min interval", 
    ylab = "Number of Steps", main = "Average number of steps taken", 
    col = "blue")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

To find that, 
```{r}
MaxStepAvg <- which.max(timeSeriesPat)
IntervalWithMaxStepAvg <- names(MaxStepAvg)
```

The 5-minute interval `r IntervalWithMaxStepAvg` contains the maximum number of steps on average across all the days.

## Inputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
NumOfNAs <- sum(is.na(activity))
```

There are `r NumOfNAs` missing values in the dataset.

Devise a strategy for filling in all of the missing values in the dataset.
```{r}
StepAvg <- aggregate(steps ~ interval, data = activity, FUN = mean)

fillNAs <- numeric()
for (i in 1:nrow(activity)) {
    obs <- activity[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepAvg, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNAs <- c(fillNAs, steps)
}
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
filledActivity <- activity
filledActivity$steps <- fillNAs
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

We shall do this with ggplot2 for a change.

```{r}
library(ggplot2)

AllStepsPerDay <- aggregate(steps ~ date, filledActivity, sum)

colnames(AllStepsPerDay) <- c("date","steps")

ggplot(AllStepsPerDay, aes(x = steps)) + 
       geom_histogram(fill = "black", binwidth = 1000) + 
        labs(title="Histogram of Steps Taken per Day", 
             x = "Number of Steps per Day", y = "Number of times in a day(Count)") + theme_bw()
```

In order to find the mean, we can just use the mean function on the new steps column
```{r}
filledMeanSteps <- mean(AllStepsPerDay$steps, na.rm=TRUE)
filledMeanSteps
filledMedianSteps <- median(AllStepsPerDay$steps, na.rm=TRUE)
filledMedianSteps
```
The new mean is `r filledMeanSteps` and the new median is `r filledMedianSteps`. From the result, we can see that the mean remained the same while the median has decreased slightly by filling in the NAs.


## Are there differences in activity patterns between weekdays and weekends?
