---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
head(activity)
```

```{r}
names(activity)
```

```{r}
library(lattice)

activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity$interval <- as.factor(activity$interval)
```

## What is mean total number of steps taken per day?

Let's use the aggregate function for this.
```{r}
StepsTaken <- aggregate(steps ~ date, data = activity, sum,na.rm = TRUE)
```

We can view a bar chart for this now
```{r}
hist(StepsTaken$steps,main = "Total number of Steps taken by Day",xlab="Steps",ylab = "Days", col="blue")
```

In order to find the mean, we can just use the mean function on the steps column
```{r}
meanSteps <- mean(StepsTaken$steps,na.rm=TRUE)
medianSteps <- median(StepsTaken$steps,na.rm=TRUE)
```


The Mean is `r round(meanSteps,1)` and the median is `r medianSteps`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
